\section*{Exam structure}
What follows is the general structure of a DEVI exam.

\paragraph{Question I: formal rules} \ \\

\textbf{General shape of the question:} \textit{given a set of rules, and given the starting point, what is the result of excution? Give the intermediate steps as well.}

\ \\ 

\textbf{Points:} \textit{25\%.}

\ \\ 

\textbf{Grading:} \textit{All points for correct answer, partial trajectories count for exactly half the score, a completely orthogonal trajectory counts for no points.}

\ \\ 

\textbf{Associated learning goals:} \texttt{LMC}.

\ \\ 

\paragraph{Question II: program state} \ \\ 

\textbf{General shape of the question:} \textit{Fill-in the program state with the values that the variables assume while running the sample below.}

\ \\ 

\textbf{Points:} \textit{25\%.}

\ \\ 

\textbf{Grading:} \textit{Full points if all stack frames and return types are correctly listed, with the right time stamps. Half points if at least half of all stack frames is listed. Zero points otherwise.}

\ \\ 

\textbf{Associated learning goals:} \texttt{CMC}.

\ \\ 

\paragraph{Question III: variables, expressions, and data types}

\ \\ 

\textbf{General shape of the question:} \textit{What is the value and the type of all variables after execution of the following code?}

\ \\ 

\textbf{Points:} \textit{25\%.}

\ \\ 

\textbf{Grading:} \textit{All values and types are correct: full-points. At least half the values and at least half the types are correct: half points. Zero points otherwise.}

\ \\ 

\textbf{Associated learning goals:} \texttt{VAR}, \texttt{EXPR}.

\ \\ 

\paragraph{Question IV: control flow}

\ \\ 

\textbf{General shape of the question:} \textit{What is the value of all variables after execution of the following code?}

\ \\ 

\textbf{Points:} \textit{25\%.}

\ \\ 

\textbf{Grading:} \textit{All values are correct: full-points. At least half the values are correct: half points. Zero points otherwise.}

\ \\ 

\textbf{Associated learning goals:} \texttt{COND}, \texttt{LOOP}.

\ \\ 
