\section*{Exam sample 2}
What follows is a concrete example of the exam.


\paragraph{Question I: formal rules} \ \\

\textit{You start at point (-10,20). Take 5 steps in the direction (5,-5). Then take steps in the direction (0, 10) until you are above (0, 10). Where do you end up?}

\ \\ 

\textbf{Answer:} \textit{The trajectory is:}

\begin{lstlisting}


\end{lstlisting}

\ \\ 

\textbf{Points:} \textit{25\%.}

\ \\ 
\ \\ 

\paragraph{Question II: program state} \ \\ 

\textit{Fill-in the program state with the values that the variables assume while running the sample below.}

\begin{lstlisting}
count = 1
for i in range(0, 5):
    count *= i
\end{lstlisting}

\ \\ 

\textbf{Answer:} \textit{The variable allocations are:}

%\begin{tabular}{| c | c | c | c | c | c | c | c | c | c | c | c |}
%\hline
%\textbf{y} & 1    & 1 & 2 & 2 & 4 & 4 & 8 & 8 & 16 & 16 & 32 \\
%\hline
%\textbf{i} & n.a. & 0 & 0 & 1 & 1 & 2 & 2 & 3 & 3  &  4 &  4 \\
%\hline
%\end{tabular}

\ \\

\textbf{Points:} \textit{25\%.}

\ \\ 

\textbf{Grading:} \textit{Full points if all values are correctly listed in the right order. Half points if at least half of values are listed in the right order. Zero points otherwise.}

\ \\ 

\textbf{Associated learning goals:} \texttt{CMC}.

\ \\ 

\paragraph{Question III: variables, expressions, and data types}

\ \\ 

\textit{What is the value and the type of all variables after execution of the following code?}
\begin{lstlisting}
h = input("Wat is je naam?")
j = "Hello {}".format(h)
k = 10 / 3
l = k <= 3 or True
i = "Hello" + 1
\end{lstlisting}

\ \\ 

\textbf{Answer:} \textit{The value and type of all variables after execution is:}

%\begin{tabular}{| l | c | c | }
%\hline
%\textbf{Variable} & \textbf{Value} & \textbf{Type} \\
%\hline
%h & 'Youri' & string \\
%\hline
%j & 'Hello Youri' & str \\
%\hline
%k & 3.333333 & float \\
%\hline
%l & True & boolean \\
%\hline
%i & Error, omdat je geen plus kan doen met een string en een int & error \\
%\hline
%\end{tabular}

\ \\ 

\textbf{Points:} \textit{25\%.}

\ \\ 

\textbf{Grading:} \textit{All values and types are correct: full-points. At least half the values and at least half the types are correct: half points. Zero points otherwise.}

\ \\ 

\textbf{Associated learning goals:} \texttt{VAR}, \texttt{EXPR}.

\ \\ 

\paragraph{Question IV: control flow}

\ \\ 

\textbf{General shape of the question:} \textit{What is the value of all variables after execution of the following code?}

\ \\ 

\textbf{Concrete example of question:} \textit{Draw what is printed on the screen after execution of the following code?}

\begin{lstlisting}
output = ""
for i in range(0, 4):
    for j in range(0, 4):
        if (i == 0 or i == 3) or (j == 0 or j == 3):
            output += "0"
        else:
            output += "="
    output += "\n"
print(output)
\end{lstlisting}

\ \\ 

\textbf{Concrete example of answer:} \textit{The screen looks like:}
\begin{lstlisting}

0000
0==0
0==0
0000

\end{lstlisting}

\ \\ 

\textbf{Points:} \textit{25\%.}

\ \\ 

\textbf{Grading:} \textit{All values are correct: full-points. At least half the values are correct: half points. Zero points otherwise.}

\ \\ 

\textbf{Associated learning goals:} \texttt{COND}, \texttt{LOOP}.

\ \\ 
