\documentclass[11pt, a4paper]{article}
\usepackage{opgave}
\usepackage{fancyStyle}

\renewcommand{\course}{DEV01-1}
\renewcommand{\titel}{Assignment 2}
\renewcommand{\auteur}{Youri Tjang}
\renewcommand{\auteurlang}{\auteur\\ \href{mailto:y.s.tjang@hr.nl}{\texttt{y.s.tjang@hr.nl}}}

\setboolean{isTentamen}{false}
\setboolean{showAntwoord}{false}

\usepackage{graphicx}

\begin{document}
\setcounter{opgave}{1}
\opgave{}{1}
Write a program using stdNt that satisfies the following requirements:

For all exercises a student will be chosen by chance will run their program in front of the class.


\subopgave{Write a program that allows any student to reach the goal.}{}{1}
\begin{itemize}
    \item You need to reach the door of the classroom.
    \item No whiles and if's
    \item Between the student and the door there are no objects
    \item the language consists only of the following instructions
    \begin{itemize}
    	\item take $x$ steps
	\item turn \textit{left/right} $\alpha$ degrees
	\item sit, stand, smile
    \end{itemize}
\end{itemize}
\subopgave{Also use the memory model (program counter, variables, etc) shown in the slides. And write a complete run of your program.}{}{1}
\subopgave{Extend your program so that you avoid objects in the room.}{}{1}

\opgave{}{1}
In exercise 1 you have written a program that lets \textbf{you} find the door. In this exercise you will make that program more generic.
\subopgave{Adjust the program so that any arbitrary student in the current room (including objects) can find the door. In order to achieve this the language from exercise 1 is extended with the following instructions:
\begin{itemize}
   \item \textit{if's} and \textit{whiles}
   \item distance to \textit{board/door/window}  $\rightarrow$  \textit{returns a \textbf{number}, which is the distance from the student to the object}
   \item collisionWithObstacle $\rightarrow$  \textit{returns \textbf{true} if there is an obstacle less than 1 step in front of the student, otherwise returns \textit{\textbf{false}}}
 \end{itemize}
    }{}{1}
\subopgave{Discuss the differences between: is it shorter, is it more readable, are the amount of variables the same, etc?}{}{1}

\einde
\end{document}